---
title: "Assignment 3"
author: "JIANG Rui 3035869637  CHAN Yat Tin 3035869651"
date: '2022-05-12'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages
```{r}
# install.packages("stargazer")
suppressMessages(library(AER))
library(stargazer)
library(hdm)
data("Affairs")
?Affairs
```

```{r}
Affairs$happy=(Affairs$rating==5)
Affairs$Cheater=(Affairs$affairs==12)
Affairs$age2=Affairs$age^2
Affairs$yearsmarried2=Affairs$yearsmarried^2
summary(Affairs)
```

# Part A
```{r}
ct <- lm(formula=Cheater~happy+gender+age+age2+children+as.factor(religiousness)+education+as.factor(occupation), data=Affairs)
a<-as.data.frame(ct$coefficients)

# Compute heteroskedasticity-robust standard errors and significance levels, where NAME is the name of a regression-type object created by lm() and other similar commands
ROBUST <- coeftest(ct,vcov=vcovHC(ct, type = "HC0"))
# ROBUST
stargazer(ROBUST, type="text",dep.var.labels="Cheater",title="Part A: Report the estimated coefficients, their heteroskedasticity-robust standard errors, and significance levels")
happy_coef <- as.character(ROBUST[2,1])
happy_se <- as.character(ROBUST[2,2])

# report the 90% confidence interval for the estimated coefficient of happy, where ROBUST is the name of the object created by the coeftest() command above
all_result <- confint(ROBUST, level=0.90)

# select happy result 
happy_df <- as.data.frame(all_result[2,])
# happy_df[2,1]
happy_90_ci <- paste("[",as.character(happy_df[1,1]),",",as.character(happy_df[2,1]),"]",sep="")
# happy_90_ci

# cbind
table <- cbind(happy_coef,happy_se,happy_90_ci)
colnames(table) <- c("happy coefficient", "happy SE", "happy 90% CI")

# present result in stargazer command 
stargazer(table, type="text", title="Part A: Summary of happy coefficients")
```

1. What is the estimated effect of happy on the probability of having regular affairs?
Ans: Being very satisfied with marriages is expected to reduce the probability of having regular affairs by 0.0315.

2. Is the estimated effect significant?
Ans: Since the p-value is 0.090841 < 0.10, the estimated effect is statistically significant at 10% significance level. 

3. Does the sign of estimated effect makes sense?
Ans: Yes. If you have a happy marriage, you are less likely to find partners outside and thus have affairs.


4. Comment on the magnitude of the effect: does it make sense?
Ans: From my perspective, the magnitude is drastically less than expected because marriage satisfaction should be one of the major factors dissuading people from having affairs.  

# Part B
```{r}
# IV
iv <- ivreg(Cheater~happy+gender+age+age2+children+as.factor(religiousness)+education+as.factor(occupation) | yearsmarried + yearsmarried2 + (yearsmarried+yearsmarried2)*( gender + age + age2 + children + as.factor(religiousness)+education+as.factor(occupation) ), data=Affairs)
iv_ROBUST <- coeftest(iv,vcov=vcovHC(iv, type = "HC0"))
# iv_ROBUST
# coefficient
# iv_ROBUST[2,1]
# SE
# iv_ROBUST[2,2]

# IV result presentation
stargazer(iv_ROBUST, type="text",dep.var.labels = "Cheater" ,title="Part B: Report the IV estimates for the coefficient on happy and its heteroskedasticity-robust standard errors and significance levels")

# CI
happy_df_b <- as.data.frame(confint(iv_ROBUST,level=0.90)[2,])
happy_90_ci_b <- paste("[",as.character(happy_df_b[1,1]),",",as.character(happy_df_b[2,1]),"]",sep="")
stargazer(happy_90_ci_b, type="text",title="Part B: Report the heteroskedasticity-robust IV-based 90% confidence interval for the coefficient of happy")

# First stage
first_stage <- lm(happy~yearsmarried + yearsmarried2 + (yearsmarried+yearsmarried2)*( gender + age + age2 + children + as.factor(religiousness)+education+as.factor(occupation)) , data=Affairs)
first_stage_robust <- coeftest(first_stage, type="HC0")
# first_stage_robust
stargazer(first_stage_robust,type="text",dep.var.labels = "Happy",title="Part B: Report the first stage estimates with their heteroskedasticity-robust standard errors and significance levels" )
```
1. Report the IV estimates for the coefficient on happy and its heteroskedasticity-robust standard errors and significance levels. 

Ans: The IV estimates are shown in the 1st table.  


2. Report the heteroskedasticity-robust IV-based 90% confidence interval for the coefficient of happy. 

Ans: The heteroskedasticity-robust IV-based 90% confidence interval is [-0.214886041453851,0.117377553751838].  

3. Comment on the IV estimates of the magnitude of the effect.

Ans: The IV estimates of the magnitude of the effect is -0.04875424, which is still much less than expected.

4. Discuss the significance of IV estimates.

Ans: Age, age2, education, and some other factor variables such as religion and occupation are statistically significant. Other variables are statistically insignificant.  

5. Compare with OLS estimates in part A.

Ans: Compared above, the magnitude of estimate on happy increases a little bit.   

6. Report the first stage estimates with their heteroskedasticity-robust standard errors and significance levels. Do the IVs appear to be related to happy?

Ans: The first stage estimates are shown as the above table. Only some IVs appear to be related to happy. Many other estimates are statitically insignificant.    


# Part c
```{r}
lasso.iv <- rlassoIV(Cheater~happy+gender+age+age2+children+as.factor(religiousness)+education+as.factor(occupation) | yearsmarried + yearsmarried2 + (yearsmarried+yearsmarried2)*( gender + age + age2 + children + as.factor(religiousness)+education+as.factor(occupation) ), data=Affairs,select.X=TRUE,select.Z=TRUE)

# post lasso estimates
summary(lasso.iv)
lasso.iv$coefficients
lasso.iv$se

# 90% CI
lasso.iv.ci <- confint(lasso.iv,level=0.90)
happy_90_ci_c <- paste("[",as.character(-0.8310902),",",as.character(-0.1418015),"]",sep="")

# Present result 
post_lasso_table <- cbind(lasso.iv$coefficients,happy_90_ci_c)
colnames(post_lasso_table) <- c("Estimate","90% Confidence Interval")
rownames(post_lasso_table) <- "happy"
stargazer(post_lasso_table, type="text", title="Report the post-Lasso IV-based estimates for the effect of happy on the probability of regular affairs and the 90% confidence interval for the estimated post-Lasso effect of happy")

# Second stage
Second <- rlasso(Cheater~happy+gender+age+age2+children+as.factor(religiousness)+education+as.factor(occupation),data=Affairs )
second_var <- which(Second$index==TRUE)
second_var 

# First stage
First <- rlasso(happy~yearsmarried + yearsmarried2 + (yearsmarried+yearsmarried2)*( gender + age + age2 + children + as.factor(religiousness)+education+as.factor(occupation) ), data=Affairs)
which(First$index==TRUE)
```
1. Report the post-Lasso IV-based estimates for the effect of happy on the probability of regular affairs.

Ans: The result is as above. 

2. Report the 90% confidence interval for the estimated post-Lasso effect of happy

Ans: The result is as above. 

3. Use rlasso() to check which controls are useful for predicting Cheater

Ans: Occupation is useful for predicting Cheater.

4. Use rlasso() to check which IVs and controls are useful for predicting happy (the first stage)

Ans: yearsmarried, childrenyes, as.factor(occupation), yearsmarried:childrenyes, yearsmarried2:as.factor(religiousness),  and yearsmarried2:as.factor(occupation) are IVs and controls useful for predicting happy.



# Part d
```{r}
happy_coef_ptd <- cbind(happy_coef, iv_ROBUST[2,1], lasso.iv$coefficients)
happy_se_ptd <- cbind(happy_se, iv_ROBUST[2,2], lasso.iv$se)
happy_ci_ptd <- cbind(happy_90_ci,happy_90_ci_b,happy_90_ci_c) 
happy_summary <- rbind(happy_coef_ptd, happy_se_ptd, happy_ci_ptd)
colnames(happy_summary) <- c("Part A OLS ", "Part B IV", "Part C RLASSO IV")
rownames(happy_summary) <- c("Happy Coefficient", "SE", "90% CI")
stargazer(happy_summary, type="text",title="Results for the effect of happy on Cheater from parts A-C")
```

1. Compare and discuss the three sets of results.

Ans: Part b and part c have much higher SE, suggesting that IV-based methods produce much noisier estimates. When we perform post-lasso, the coefficient on happy increases significantly in absolute value. Yet the 90% CI for RLASSO IV is much wider compared that of the other two parts.   

2. Which estimates make most sense?

Ans: Part C as happy should not be accounting for such minor effects on the probability of having affairs as in part a and part b, though the high SE is the trade-off we have to admit.  

3. Are Lasso/post-Lasso methods useful here for improving estimation results?

Ans: Yes. The estimates increase significantly in scale. Yet, standard error as well as wider CI are the trade-offs.
