---
title: "Assignment2"
author: "JIANG Rui CHAN Yat Tin"
date: "4/30/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install packages 
```{r}
install.packages("carData")
install.packages("glmnetUtils")
install.packages("margins")
```

## Load packages 
```{r}
library(carData)
library(glmnetUtils)
library(margins)
library(glmnetUtils)
```

## Dataset Info
The data set contains information on ex-convicts collected during a period of 52 weeks after their release:

1. arrest 1 if arrested.

2. fin yes if received financial aid.

3. Additional variables on age, marital status, work experience, education, and etc.

4. week The week of first arrest after release. 52 if not arrested during the period.

5. The variables emp1–emp52 track their employment status by week with value yes if employed.

Generate a new variable Wks that measures the fraction of time an ex-convict was unemployed before re-arrested or until the end of the 52 weeks period:
```{r}
?Rossi
Rossi_df <- as.data.frame(Rossi)
summary(Rossi)
Rossi$Weeks_worked=0
for (i in 1:nrow(Rossi)){
  last_week_col=Rossi$week[i]+10 #last week before arrest column number
  for (j in 11:last_week_col){
  Rossi$Weeks_worked[i]=Rossi$Weeks_worked[i]+(Rossi[i,j]=="yes")
  }
}
Rossi$Wks=(Rossi$week-Rossi$Weeks_worked)/Rossi$week
```

Part A -- Estimate the following logit model using glm():
1. Report the estimattion results using the summary() command.
2. Report the AME for fin including the 95% confidence interval. Does financial aid have a significant effect on re-arrests?
```{r}
Logit <- glm(arrest~fin+age+race+wexp+mar+paro+prio+educ+Wks,data=Rossi,family=binomial(link="logit"))
coef(summary(Logit))
summary(margins(Logit,variables="fin"))
```
 

Since the p-value is 0.0925 > 0.05, financial aid does not have a significant effect on re-arrests.
 
Part B
```{r}
# Use glmnet() to calculate adaptive LASSO
Lasso <- glmnet(arrest~fin+age+race+wexp+mar+paro+prio+educ+Wks,data=Rossi,family="binomial",alpha=0,lambda=0,standardize=FALSE, intercept=FALSE)

# Focus on finyes 
coef(Lasso)

# Generate the penalty weights
w=1/abs(coef(Lasso)[-1])
w[1:2]=0
w
```
 
Part C 
```{r}
set.seed(6083,sample.kind="Rejection")
CV_AL=cv.glmnet(arrest~fin+age+race+wexp+mar+paro+prio+educ+Wks,data=Rossi,family="binomial",alpha=0,standardize=FALSE, intercept=FALSE,penalty.factor=w)
coef(CV_AL,CV_AL$lambda.1se)
```
 
Part D
```{r}
X=model.matrix(arrest~fin+age+race+wexp+mar+paro+prio+educ+Wks,data=Rossi)[,-1]
y=Rossi$arrest

# Adjust w
adjust_w <- w
adjust_w <- w[c(2,3,4,5,6,8,10,11,12)]
adjust_w 

CV=cv.glmnet(X,y,family="binomial",alpha=0,standardize=FALSE, penalty.factor=adjust_w, nfolds=10)
plot(CV)
coef(CV,s=CV$lambda.1se)
```
Part E
```{r}
finyes <- Rossi$fin[Rossi$fin == "yes"]
Logit_E <- glm(arrest~finyes+age+raceother+wexpyes+`marnot married`+paroyes+prio+educ+Wks,data=Rossi,family=binomial(link="logit"))
coef(summary(Logit_E))
summary(margins(Logit_E,variables="fin"))
```
 
