---
title: "Assignment1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1 
Data Generating Process
```{r pressure, echo=FALSE}
dgp_1 <- function(n, k){
  # Parameter
  # 1. n -- number of observations 
  # 2. k -- number of controls 
  
  # Generate Di
  ## Generate Xij
  matrix_x <- matrix(rnorm(n*k, mean=0,sd=1), n, k)
  ## Generate Xij mean
  matrix_x_mean <- as.matrix(rowMeans(matrix_x))
  ## Generate Vi
  matrix_v <- matrix(rnorm(n*1, mean=0,sd=0.01), n, 1)
  ## Generate Di
  matrix_d <- matrix_x_mean + matrix_v
  
  # Generate Ui
  matrix_u <- matrix(rnorm(n*1, mean=0,sd=1), n, 1)
  
  # Generate Yi
  matrix_y <- matrix_d + matrix_u
  
  # Rename matrix columns 
  colnames(matrix_y) <- "yi"
  colnames(matrix_d) <- "di"
  colnames(matrix_x) <- c("xi,1","xi,2","xi,3","xi,4","xi,5",
                          "xi,6","xi,7","xi,8","xi,9","xi,10",
                          "xi,11","xi,12","xi,13","xi,14","xi,15",
                          "xi,16","xi,17","xi,18","xi,19","xi,20")
  colnames(matrix_u) <- "ui"
  
  # Column bind the matrixes 
  dataset <- cbind(matrix_y, matrix_d, matrix_x, matrix_u)
  
  # Convert matrix to dataframe 
  dataset.data.frame <- as.data.frame(dataset)
  
  return(dataset.data.frame)
}


```

Part a -- Generate n=100 observations on the dependent variable Yi, the main regressor Di, and k=20 controls/covariates Xi,j,j=1,…,k, and then save it into a dataframe 
```{r}
dataset.data.frame_1 <- dgp_1(100, 20)
```

Part b -- Using lm(), estimate the regression of Yi against Di only and report the standard error and p-value on the coefficient of Di
```{r}
reg_b <- lm(yi~di, data=dataset.data.frame_1)
summary(reg_b)
```

Part c -- Using lm(), estimate the regression of Yi against Di with all 20 controls included and report the standard error and p-value on the coefficient of Di
Compare with the standard error and p-value in part (b). Discuss the differences and if they make sense
```{r}
reg_c= lm(yi~di+`xi,1`+`xi,2`+`xi,3`+`xi,4`+`xi,5`+`xi,6`+`xi,7`+`xi,8`+`xi,9`+`xi,10`+`xi,11`+`xi,12`+`xi,13`+`xi,14`+`xi,15`+`xi,16`+`xi,17`+`xi,18`+`xi,19`+`xi,20`, data=dataset.data.frame_1)
summary(reg_c)
```

Part d -- MC simulations
Repeat parts (a)-(c) R=1000 times 
1. Each time generating a new data set
2. Store the results for the standard errors and p-values obtained in each repetition.
3. Report the average standard errors (across the MC repetitions) for the OLS estimator on Di when the controls are excluded and included.
4. Report the fraction of MC repetitions in which the OLS estimators are significant at the 5% signifincance level.
5. Compare the results (for when the controls are excluded and included) and discuss the differences.

```{r}
# SE and P-value of Di coefficient trial -- test to see the location of SE and P-value in the regression output table for each regression 
reg_b_trial <- summary(reg_b)$coefficients
reg_b_trial
reg_b_trial[2,2] # SE without control
reg_b_trial[2,4] # P-value without control 
reg_c_trial <- summary(reg_c)$coefficients
reg_c_trial[2,2] # SE with control
reg_c_trial[2,4] # P-value with control 
```

```{r}
# Create simulated values that are reproducible

# Create a empty dataframe storing SE and p-value of the coefficient of Di 
result_frame <- data.frame(matrix(ncol=4, nrow=0))
colnames(result_frame) <- c("SE without control", "P-value without control", "SE with control", "SE without control")

# Use for loop to do part a to c for 1000 times 
for (i in 1:1000){
  # In each time, we generate a dataset 
  dataset.data.frame_2 <- dgp_1(100,20)
  
  # Redo part b
  reg_b_d = lm(yi~di, data=dataset.data.frame_2)
  reg_b_d_df = summary(reg_b_d)$coefficients
  
  # Extract SE and P-value without control
  reg_b_d_df[2,2] # SE without control
  reg_b_d_df[2,4] # P-value without control 
  
  # Redo part c
  reg_c_d= lm(yi~di+`xi,1`+`xi,2`+`xi,3`+`xi,4`+`xi,5`+`xi,6`+`xi,7`+`xi,8`+`xi,9`+`xi,10`+`xi,11`+`xi,12`+`xi,13`+`xi,14`+`xi,15`+`xi,16`+`xi,17`+`xi,18`+`xi,19`+`xi,20`, data=dataset.data.frame_2)
  reg_c_d_df = summary(reg_c_d)$coefficients
  
  # Extract SE and P-value with control
  reg_c_d_df[2,2] # SE with control
  reg_c_d_df[2,4] # P-value with control
  
  # Add SE and p-values with the previous df in a row-wise manner
  result_frame[nrow(result_frame)+1,] <- c(reg_b_d_df[2,2],reg_b_d_df[2,4],reg_c_d_df[2,2],reg_c_d_df[2,4])
  
}
```









# Question 2 -- Data Generating Process (DGP)
Part a -- Define a custom function that generates n observations from the following data generating process (DGP)
```{r}
dgp_2 <- function(n, l, a, p){
  # Parameter:
  # 1. n -- number of observations 
  # 2. l -- number of instruments 
  # 3. a -- coefficient of Di in Yi=αDi+Ui
  # 4. p -- coefficient of Ui =ρVi+ϵi
  
  # Generate Di
  ## Generate Zi,1
  n = 100
  l = 40
  a = 2 
  p = 10
  matrix_z_i1 = matrix(rnorm(n*1, mean=0,sd=1), n, 1)
  
  ## Generate 1/[(n)^0.5]Sigma(Zi,j)
  ### Generate Zi,j
  matrix_z_ij = matrix(rnorm(n*(l-1), mean=0,sd=1), n, l-1)
  ### Perform summation in each row 
  matrix_z_ij_sum = as.matrix(rowSums(matrix_z_ij))
  matrix_z_ij_sum_final =  (matrix_z_ij_sum) / n^0.5
  
  ## Generate Vi 
  matrix_v_i = matrix(rnorm(n*1, mean=0,sd=1), n, 1)
  ## Generate Di
  matrix_d_i = matrix_z_i1 + matrix_z_ij_sum_final + matrix_v_i
  
  # Generate Ui
  ## Generate epsilon i
  matrix_e_i = matrix(rnorm(n*1, mean=0,sd=1), n, 1)
  ## Generate Ui
  matrix_u_i = p * matrix_v_i +  matrix_e_i
  
  # Generate Yi
  matrix_y_i = 2 * matrix_d_i + matrix_u_i
  
  # Rename matrix columns 
  colnames(matrix_z_i1) <- "zi,1"
  colnames(matrix_z_ij) <- c("zi,2","zi,3","zi,4","zi,5",
                          "zi,6","zi,7","zi,8","zi,9","zi,10",
                          "zi,11","zi,12","zi,13","zi,14","zi,15",
                          "zi,16","zi,17","zi,18","zi,19","zi,20",
                          "zi,21","zi,22","zi,23","zi,24","zi,25",
                          "zi,26","zi,27","zi,28","zi,29","zi,30",
                          "zi,31","zi,32","zi,33","zi,34","zi,35",
                          "zi,36","zi,37","zi,38","zi,39","zi,40")
  colnames(matrix_z_ij_sum) <- "zij_sum"
  colnames(matrix_z_ij_sum_final) <- "zij_sum_final"
  colnames(matrix_v_i) <- "vi"
  colnames(matrix_d_i) <- "di"
  colnames(matrix_u_i) <- "ui"
  colnames(matrix_y_i) <- "yi"
  
  # Column bind the matrixes 
  dataset <- cbind(matrix_z_i1, matrix_z_ij, matrix_z_ij_sum, matrix_z_ij_sum_final, matrix_v_i, matrix_d_i,matrix_u_i,matrix_y_i)
  
  # Convert matrix to dataframe 
  dataset.data.frame <- as.data.frame(dataset)
  
  return(dataset.data.frame)

  
}
  
```


trial 
```{r}
data.frame.iv <- dgp_2(100, 40, 2, 10)
```

Part b 
1. Perform R=103 independent MC repetitions, each time generating a new set of data using the DGP in (a) with n=100 and ℓ=40.
2. At each MC repetition, compute the 2SLS estimators for the IV regression of Yi against Di using the following IVs:
    Only the first IV Zi,1.
    The first 5 IVs Zi,1,…,Zi,5.
    The first 10 IVs.
    The first 15 IVs.
    The first 20 IVs.
    The first 25 IVs.
    The first 30 IVs.
    The first 35 IVs.
    All 40 IVs.
3. Save estimates for all the 9 2SLS estimators described above obtained at each MC repetition.
4. Compute the simulated bias
5. Plot the absolute value of the simulated bias of the 2SLS estimator (on the vertical axis) against the number of IVs used (on the horizontal axis).
6. Discuss the effect of the number of IVs on the simulated bias of the 2SLS. Does using many IVs result in substantial bias?

Test 
```{r}
library(AER)
# Create a empty dataframe storing estimator of Di 
iv_result_frame <- data.frame(matrix(ncol=9, nrow=0))
colnames(iv_result_frame) <- c("Only the first IV Zi,1", "The first 5 IVs Zi,1,…,Zi,5", "The first 10 IVs", "The first 15 IVs","The first 20 IVs", "The first 25 IVs", "The first 30 IVs", "The first 35 IVs","All 40 IVs")

IVmodel <- ivreg(yi~di | data.frame.iv[,1:1],data=data.frame.iv)
iv_di_coef <- summary(IVmodel)$coefficients 
di_1 <- iv_di_coef[2,2]

matrix_df <- as.matrix(data.frame.iv)

IVmodel <- ivreg(yi~di | matrix_df[,1:5],data=data.frame.iv)
iv_di_coef <- summary(IVmodel)$coefficients 
di_2 <- iv_di_coef[2,2]

IVmodel <- ivreg(yi~di | matrix_df[,1:10],data=data.frame.iv)
iv_di_coef <- summary(IVmodel)$coefficients 
di_3 <- iv_di_coef[2,2]

IVmodel <- ivreg(yi~di | matrix_df[,1:15],data=data.frame.iv)
iv_di_coef <- summary(IVmodel)$coefficients 
di_4 <- iv_di_coef[2,2]

IVmodel <- ivreg(yi~di | matrix_df[,1:20],data=data.frame.iv)
iv_di_coef <- summary(IVmodel)$coefficients 
di_5 <- iv_di_coef[2,2]

IVmodel <- ivreg(yi~di | matrix_df[,1:25],data=data.frame.iv)
iv_di_coef <- summary(IVmodel)$coefficients 
di_6 <- iv_di_coef[2,2]

IVmodel <- ivreg(yi~di | matrix_df[,1:30],data=data.frame.iv)
iv_di_coef <- summary(IVmodel)$coefficients 
di_7 <- iv_di_coef[2,2]

IVmodel <- ivreg(yi~di | matrix_df[,1:35],data=data.frame.iv)
iv_di_coef <- summary(IVmodel)$coefficients 
di_8 <- iv_di_coef[2,2]

IVmodel <- ivreg(yi~di | matrix_df[,1:40],data=data.frame.iv)
iv_di_coef <- summary(IVmodel)$coefficients 
di_9 <- iv_di_coef[2,2]

# Add to dataframe 
iv_result_frame[nrow(iv_result_frame)+1,] <- c(di_1,di_2,di_3,di_4,di_5,di_6,di_7,di_8,di_9)
iv_result_frame
```
Run 1000 times and save estimates 
```{r}
# Create a empty dataframe storing estimator of Di 
iv_result_frame <- data.frame(matrix(ncol = 9, nrow = 0))
colnames(iv_result_frame) <- c(
  "Only the first IV Zi,1",
  "The first 5 IVs Zi,1,…,Zi,5",
  "The first 10 IVs",
  "The first 15 IVs",
  "The first 20 IVs",
  "The first 25 IVs",
  "The first 30 IVs",
  "The first 35 IVs",
  "All 40 IVs"
)
for(i in 1:1000){
  # Generate dataset 
  data.frame.iv <- dgp_2(100, 40, 2, 10)
  
  IVmodel <- ivreg(yi ~ di | data.frame.iv[, 1:1], data = data.frame.iv)
  iv_di_coef <- summary(IVmodel)$coefficients
  di_1 <- iv_di_coef[2, 2]
  
  matrix_df <- as.matrix(data.frame.iv)
  
  IVmodel <- ivreg(yi ~ di | matrix_df[, 1:5], data = data.frame.iv)
  iv_di_coef <- summary(IVmodel)$coefficients
  di_2 <- iv_di_coef[2, 2]
  
  IVmodel <- ivreg(yi ~ di | matrix_df[, 1:10], data = data.frame.iv)
  iv_di_coef <- summary(IVmodel)$coefficients
  di_3 <- iv_di_coef[2, 2]
  
  IVmodel <- ivreg(yi ~ di | matrix_df[, 1:15], data = data.frame.iv)
  iv_di_coef <- summary(IVmodel)$coefficients
  di_4 <- iv_di_coef[2, 2]
  
  IVmodel <- ivreg(yi ~ di | matrix_df[, 1:20], data = data.frame.iv)
  iv_di_coef <- summary(IVmodel)$coefficients
  di_5 <- iv_di_coef[2, 2]
  
  IVmodel <- ivreg(yi ~ di | matrix_df[, 1:25], data = data.frame.iv)
  iv_di_coef <- summary(IVmodel)$coefficients
  di_6 <- iv_di_coef[2, 2]
  
  IVmodel <- ivreg(yi ~ di | matrix_df[, 1:30], data = data.frame.iv)
  iv_di_coef <- summary(IVmodel)$coefficients
  di_7 <- iv_di_coef[2, 2]
  
  IVmodel <- ivreg(yi ~ di | matrix_df[, 1:35], data = data.frame.iv)
  iv_di_coef <- summary(IVmodel)$coefficients
  di_8 <- iv_di_coef[2, 2]
  
  IVmodel <- ivreg(yi ~ di | matrix_df[, 1:40], data = data.frame.iv)
  iv_di_coef <- summary(IVmodel)$coefficients
  di_9 <- iv_di_coef[2, 2]
  
  # Add to dataframe
  iv_result_frame[nrow(iv_result_frame) + 1, ] <-
    c(di_1, di_2, di_3, di_4, di_5, di_6, di_7, di_8, di_9)
  
}
```

Calculate simulated bias 
```{r}
# Generate new column -- a(hat) - a(=2 in this case )
# for(i in 1:ncol(iv_result_frame)){
#   iv_result_frame[paste(as.character( (i-1)* 5), "Bias")] <- iv_result_frame[,i] - 2
# }
iv_result_frame["1stBias"] <- iv_result_frame["Only the first IV Zi,1"] - 2
iv_result_frame["5thBias"] <- iv_result_frame["The first 5 IVs Zi,1,…,Zi,5"] -2
iv_result_frame["10thBias"] <- iv_result_frame["The first 10 IVs"] - 2
iv_result_frame["15thBias"] <- iv_result_frame["The first 20 IVs"] -2
iv_result_frame["20thBias"] <- iv_result_frame["The first 20 IVs"] -2
iv_result_frame["25thBias"] <- iv_result_frame["The first 25 IVs"] -2
iv_result_frame["30thBias"] <- iv_result_frame["The first 30 IVs"] -2
iv_result_frame["35thBias"] <- iv_result_frame["The first 35 IVs"] -2
iv_result_frame["40thBias"] <- iv_result_frame["All 40 IVs"] -2
```

```{r}
# Calculate means of bias separately for each column and create a new column to store simulated bias
for (i in 1:1000){
  iv_result_frame[i,"1stOverallBias"] = mean(iv_result_frame$`1stBias`[i:1000])
  iv_result_frame[i,"5thOverallBias"] = mean(iv_result_frame$`5thBias`[i:1000])
  iv_result_frame[i,"10thOverallBias"] = mean(iv_result_frame$`10thBias`[i:1000])
  iv_result_frame[i,"15thOverallBias"] = mean(iv_result_frame$`15thBias`[i:1000])
  iv_result_frame[i,"20thOverallBias"] = mean(iv_result_frame$`20thBias`[i:1000])
  iv_result_frame[i,"25thOverallBias"] = mean(iv_result_frame$`25thBias`[i:1000])
  iv_result_frame[i,"30thOverallBias"] = mean(iv_result_frame$`30thBias`[i:1000])
  iv_result_frame[i,"35thOverallBias"] = mean(iv_result_frame$`35thBias`[i:1000])
  iv_result_frame[i,"40thOverallBias"] = mean(iv_result_frame$`40thBias`[i:1000])

}

```

