---
title: "Post-Lasso and double Lasso: growth rate example"
output: html_document
---


```{r}
library(hdm)
```

### The model
$$\begin{equation}
(\Delta\log GDP)_i=\alpha \cdot GDP^0_i+U_i.
\end{equation}$$

* $(\Delta\log GDP)_i$ is the change in the log of GDP per capita of country $i$ between periods $t_1$ and $t_2$, $t_2>t_1$.
* $GDP^0_i$ is the GDP per capita of country $i$ in period $t$
* **The Catching Up Hypothesis**: $\alpha <0$. The rate of growth slows down, less developed countries catch up with more developed. 
```{r}
data("GrowthData")
?GrowthData
names(GrowthData)
```

#### The hypothesis fails:
```{r}
summary(lm(Outcome~gdpsh465,data=GrowthData))
```

### An alternative model:

* Institutions and technology matter. 
* The catching up hypothesis works with similar countries only.
* Need to control for the characteristics.

$$\begin{equation}
(\Delta\log GDP)_i=\alpha \cdot GDP^0_i+X_i'\beta+U_i.
\end{equation}$$


* $X_i$ is the vector of controls describing the economic conditions of country $i$ in period $t_1$.

There are a lot of potential controls:
```{r}
dim(GrowthData)
```

Let's set up estimation
```{r}
names(GrowthData)
```
```{r}
y=as.vector(GrowthData$Outcome)
D=as.vector(GrowthData$gdpsh465)
Controls=as.matrix(GrowthData)[,-c(1,2,3)]
```

* `-c(1,2,3)` instructs to exclude the first 3 variables in `GrowthData`: 
  * `Outcome`   
  * `intercept`
  * `gdpsh465`  

OLS regression with all controls:
```{r}
Full=lm(y~D+Controls)
head(coef(summary(Full)),2)
```

* The estimate is negative but the std.err is too large - too many controls.

### Post-Lasso with Double Lasso

```{r}
?rlassoEffect
```


```{r}
Effect<-rlassoEffect(Controls,y,D,method="double selection")
summary(Effect)
```

* A negative significant estimate!

Included controls:
```{r}
sum(Effect$selection.index==TRUE)
Effect$selection.index[Effect$selection.index==TRUE]
```
```{r}
sum(Effect$selection.index==TRUE)
Effect$selection.index[Effect$selection.index==TRUE]
```

#### Double Lasso selected 7 controls:

* `bmp1l`: Log of the black market premium.
* `freetar`: Measure of tariff restrictions.
* `hm65`: Male gross enrollment ratio for higher education in 1965.
* `sf65`: Female gross enrollment ratio for secondary education in 1965.
* `lifee065`: Life expectancy at 0 in 1965.
* `humanf65`: Average schooling years in the female population over age 25 in 1965.
* `pop6565`: Population Proportion over 65 in 1965.


### Using the partialling out approach
```{r}
Effect_PO<-rlassoEffect(Controls,y,D,method="partialling out")
summary(Effect_PO)
```
```{r}
sum(Effect_PO$selection.index==TRUE)
Effect_PO$selection.index[Effect_PO$selection.index==TRUE]
```

